---
title: "Assignment 3"
output:
  word_document: default
  pdf_document: default
date: "2023-04-14"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ISLR)
library(glmnet)
library(caret)
library(dplyr)
library(kernlab)
library(neuralnet)
```

#Filtering variables for current project:
```{r}
carseats <- Carseats %>% select("Sales", "Price", "Advertising","Population","Age","Income","Education")
```

#QB1.Build a linear SVM regression model to predict Sales based on all other attributes
```{r}
set.seed(123)
model_1<- train(Sales~.,data=carseats, method= "svmLinear")
print(model_1)


#The values of C is "1" and R-squared values is 0.3462025. 
```

#QB2. Customize the search grid by checking the model’s performance for C parameter of 0.1,.5,1 and 10 using 2 repeats of 5-fold cross validation.
```{r}
set.seed(123)
model_2 <- trainControl(method = "repeatedcv", number = 5, repeats = 2)
model_3 <- expand.grid(C = c(0.1, 0.5, 1, 10))

model_4<- train(Sales~., data = carseats, method = "svmLinear", trControl= model_2, preProcess = c("center", "scale"), tuneGrid = model_3, tuneLength = 10)  
model_4

#The best C for combination of 0.1, 0.5, 1, 10 is 0.5, R-squared being 0.3466721.
```


#QB3.Train a neural network model to predict Sales based on all other attributes
```{r}
set.seed(123)
model_6<- train(Sales~., data = carseats, method = "nnet", preProcess = c("center", "scale"),linout = TRUE, trace = FALSE)
model_6
```

#QB4.What will be the estimated Sales for this record using the above neuralnet model? 
```{r}
new_carseats<- data.frame("Price" = 6.54, "Population" = 124, "Advertising" = 0, "Age" = 76, "Income" = 110, "Education" = 10)

model_7<- predict(model_6, new_carseats)
model_7

#The estimated Sales for current data frame, using nnet model would be 11.46014.
```





