---
title: "Assignment 2"
output:
  word_document: default
  pdf_document: default
date: "2023-04-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ISLR)
library(glmnet)
library(caret)
library(dplyr)
library(rpart)
library(rattle)
library(rpart.plot)
library(randomForest)
```


#Loading and selecting required columns from ISLR carseats dataset.
```{r}
carseats <- Carseats %>% select("Sales", "Price", "Advertising","Population","Age","Income","Education")
```

#1.Predicting Sales based on all attributes.
```{r}
model_1<- rpart(Sales~., data = carseats, method = "anova")
print(model_1)
rpart.plot(model_1)

#The attributes used at the top of Tree are Price, Advertising, Age.
```

#2.What will be the estimated Sales for this record using the decision tree model?
```{r}
new_carseats<- data.frame("Sales" = 9, "Price" = 6.54, "Population" = 124, "Advertising" = 0, "Age" = 76, "Income" = 110, "Education" = 10)
model_2<- predict(model_1, new_carseats, method = "anova")
model_2

#The estimated Sales for this record is predicted by the Decision Tree model is 9.58625.
```


#3.Using Caret function to train a random forest model.
```{r}
set.seed(123)
model_3<- train(Sales~.,data=carseats, method= "rf")
print(model_3)

#The best performance of mtry value when using method ="rf"  is 2, as the RMSE is 2.404492.
```


#4.Customizing the search grid by checking the model's performance for mtry values of 2, 3, 5 and 3-repeats of 5-fold cross validation.
```{r}
set.seed(123)
model_4 <- trainControl(method = "repeatedcv", number = 5, repeats = 3,search 
= "grid")
model_5 <- train(Sales~., data = carseats, method = "rf",trControl = model_4,tuneGrid = expand.grid(mtry=c(2,3,5)))
model_5

##The best performance of mtry value when using method ="rf"  is 2,3,5 is 3, as the RMSE is 2.401365.
```

